---
title: Alpha Diversity, PCAs, and Linear Modelling of All Three Lactations with DIM
  and Milk Volume
author: "Joseph Skarlupka V"
date: "2024-10-31"
output: html_document
---

#Loading Libraries
```{r, Libraries}
library("miLineage")
library("Maaslin2")
library("ggplot2")
library("phyloseq")
library("vegan")
library("DESeq2")
library("reltools")
library("phylosmith")
library("gridExtra")
library("ggpubr")


set.seed(81471438)
```


#Importing files of interest
```{r, import the files of interest}
#Load in the phyloseq object that contains
  #>4000 sequences

load("Physeq Object MilkProduction Paper All Samples No Rare or Abund.RData")


#Removing samples with less than 7000 sequences
physeq <- prune_samples(sample_sums(physeq) >= 7000, physeq)


okabe_ito_colors <- c("#E69F00", "#56B4E9", "#009E73")

```

#Milk Production - Creating High and Low Tiers of Milk Production Based on the top/bottom 20%
```{r}
Milk7d_T1 <- sample_data(physeq)$Milk7d_T1
Milk7d_T1 <- Milk7d_T1[order(Milk7d_T1)]
Milk7d_T1

cutoff = round((0.2 * length(Milk7d_T1)), digits=0)

#List is sorted from lowest to highest
Milk7d_T1[cutoff]
  # 65.2
#Anything lower than 65.2 is considered Low Production

Milk7d_T1[(length(Milk7d_T1) - cutoff)]
  # 110
#Anything higher than 110 is considered High Production

sample_data(physeq)$Milk_Tier <- ifelse(sample_data(physeq)$Milk7d_T1 <= 65.2, "Low", ifelse(sample_data(physeq)$Milk7d_T1 >= 110, "High", "Mid"))


rm(list=c("Milk7d_T1", "cutoff"))


```

##Adding categories for period of lactation and combined parity/period
```{r}
sample_data(physeq)$PARITY <- gsub(" ", "", sample_data(physeq)$PARITY)

sample_data(physeq)$PARITY <- as.numeric(sample_data(physeq)$PARITY)
sample_data(physeq)$DIM <- as.numeric(sample_data(physeq)$DIM)
sample_data(physeq)$PARITY <- as.numeric(sample_data(physeq)$PARITY)


sample_data(physeq)$STAGE <- ifelse(sample_data(physeq)$DIM < 101, "Early", ifelse((sample_data(physeq)$DIM >=101 & sample_data(physeq)$DIM < 202), "Middle", "Late"))

sample_data(physeq)$PARITY.GROUP <- ifelse(sample_data(physeq)$PARITY == 1, "First", ifelse(sample_data(physeq)$PARITY == 2, "Second", ifelse(sample_data(physeq)$PARITY == 3, "Third", "Multi")))

sample_data(physeq)$Parity.Stage <- paste0(sample_data(physeq)$PARITY.GROUP,".",sample_data(physeq)$STAGE)


table(sample_data(physeq)$PARITY.GROUP)
table(sample_data(physeq)$PARITY)
table(sample_data(physeq)$Parity.Stage)

```


#Rarefying the phyloseq object
```{r}
rowSums(otu_table(physeq))

#Rarefying
physeq.swab1.rare <- rarefy_even_depth(physeq, sample.size=7000)
table(sample_data(physeq.swab1.rare)$parity.group)
table(sample_data(physeq.swab1.rare)$Sample.Type_GFE.Efficiency)
table(sample_data(physeq.swab1.rare)$Sample.Type_RFI.Efficiency)

#7000 - 30 samples removed

tab <- otu_table(physeq.swab1.rare)
class(tab) <- "matrix"
#rarecurve(tab, step=50, cex=0.5, xlim=c(0,15000), ylim=c(0,2000))
rm(tab)

#Pulling back out the rarefied ASV table with new column names
ASV1 = as(otu_table(physeq.swab1.rare), "matrix")
if(taxa_are_rows(physeq.swab1.rare)){ASV1 <- t(ASV1)}
# Coerce to data.frame
ASV.7000 = as.data.frame(ASV1)
rm(ASV1)


#Pulling back out the metadata table with new column names
meta.7000 = as(sample_data(physeq.swab1.rare), "matrix")
# Coerce to data.frame
meta.7000 = as.data.frame(meta.7000)


#Pulling out new taxonomy file
tax.7000 = as(tax_table(physeq.swab1.rare), "matrix")
tax.7000 = as.data.frame(tax.7000)

```


##Abundance Cutoff of Phyloseq Object
```{r}
#Removing ASVs from which there are no samples with more than 10 sequences
physeq.swab1.abund <- prune_taxa(taxa_sums(physeq.swab1.rare)>10,physeq.swab1.rare)


```


##Pulling out the ASV, Meta, and Taxonomy of the rarefied and abundance cutoff phyloseq object
```{r}
#Pulling back out the rarefied ASV table with new column names
ASV1 = as(otu_table(physeq.swab1.abund), "matrix")
if(taxa_are_rows(physeq.swab1.abund)){ASV1 <- t(ASV1)}
# Coerce to data.frame
ASV.rare.abund = as.data.frame(ASV1)
rm(ASV1)


#Pulling back out the metadata table with new column names
meta.rare.abund = as(sample_data(physeq.swab1.abund), "matrix")
# Coerce to data.frame
meta.rare.abund = as.data.frame(meta.rare.abund)


#Pulling out new taxonomy file
tax.rare.abund = as(tax_table(physeq.swab1.abund), "matrix")
tax.rare.abund = as.data.frame(tax.rare.abund)

```


##Creating a phyloseq object of that includes an abundance cutoff, but no rarefaction
```{r}
#Removing ASVs from which there are no samples with more than 10 sequences

physeq.norare.abund <- prune_taxa(taxa_sums(physeq)>10,physeq)

save(physeq.norare.abund, file="NO RARE 1st Lactation ARS Swabs Abund Cutoff Physeq Object.RData")

```


##Adding alpha diversity metrics in phyloseq
```{r, adding alpha diversity metrics in phyloseq, then pulling out the component tables again so I can do some work in vegan}
#putting all metrics into a file
div <- phyloseq::estimate_richness(physeq.swab1.abund, measures=c("Observed", "Chao1", "ACE", "Shannon", "InvSimpson", "Simpson"))
row.names(div) <- gsub(".", "-", row.names(div), fixed=TRUE)
physeq.swab1.abund <- merge_phyloseq(physeq.swab1.abund, sample_data(div))

rm(list=c("div"))

```


##Saving the Physeq object
```{r}

save(physeq.swab1.abund, file="1st Lactation ARS Swabs Norm and Abund Physeq Object.RData")


```


#Looking at the distribution of DIM and milk production values
```{r}
ggplot(meta.rare.abund, aes(x=DIM, y=Milk7d_T1)) + 
  geom_point()+
  geom_smooth(method=lm)

shapiro.test(meta.rare.abund$DIM) # => p = 0.1229
shapiro.test(meta.rare.abund$Milk7d_T1) # => p = 0.09

ggqqplot(meta.rare.abund$DIM, ylab = "DIM")
ggqqplot(meta.rare.abund$Milk7d_T1, ylab = "Milk7d_T1")

res <- cor.test(meta.rare.abund$DIM, meta.rare.abund$Milk7d_T1, 
                    method = "spearman")
res


```

```{r}
meta <- as.matrix(sample_data(physeq))
meta <- as.data.frame(meta)
```

##FIRST LACTATION Animals
```{r}
physeq.first <- subset_samples(physeq, parity.group == "First")

Milk7d_T1 <- sample_data(physeq.first)$Milk7d_T1
Milk7d_T1 <- Milk7d_T1[order(Milk7d_T1)]
Milk7d_T1

cutoff = round((0.2 * length(Milk7d_T1)), digits=0)

#List is sorted from lowest to highest
Milk7d_T1[cutoff]
  # 61
#Anything lower than 61 is considered Low Production

Milk7d_T1[(length(Milk7d_T1) - cutoff)]
  # 87.5
#Anything higher than 87.5 is considered High Production

sample_data(physeq.first)$Milk_Tier <- ifelse(sample_data(physeq.first)$Milk7d_T1 <= 61, "Low", ifelse(sample_data(physeq.first)$Milk7d_T1 >= 87.5, "High", "Mid"))


rm(list=c("Milk7d_T1", "cutoff"))


#Applying PCA to the centered log-ratio (CLR) transformed sequence counts
ps_clr <- microbiome::transform(physeq.first, "clr")

#PCA via phyloseq
ord_clr <- phyloseq::ordinate(ps_clr, "RDA")

#Plot scree plot
phyloseq::plot_scree(ord_clr) +
  geom_bar(stat="identity", fill="blue") +
  labs(x = "\nAxis", y = "Proportion of Variance\n")

#Examine eigenvalues and % proportional variance explained
head(ord_clr$CA$eig)

sapply(ord_clr$CA$eig[1:5], function(x) x / sum(ord_clr$CA$eig))     

#Scale axes and plot ordination
clr1 <- ord_clr$CA$eig[1] / sum(ord_clr$CA$eig)
clr2 <- ord_clr$CA$eig[2] / sum(ord_clr$CA$eig)
z_first <- phyloseq::plot_ordination(physeq.first, ord_clr, type="samples", color="Milk_Tier") + 
  geom_point(size = 2) +
  coord_fixed() +  
  ylim(-12,12) +
  xlim(-12,12) +
  stat_ellipse(aes(group = Milk_Tier), linetype = 2) +
  guides(color = guide_legend(title = "Milk Volume")) +
  ggtitle("First Lactation PCA by Milk Volume Tier") +
  scale_color_manual(values = okabe_ito_colors)

z_first

ordination_plot <- plot_ordination(physeq.first, ord_clr, type="samples", color="Milk_Tier")
plot_data <- as.data.frame(ordination_plot$data)


#Test for differences between groups
#Generate distance matrix
clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean") 
#ADONIS test
vegan::adonis2(clr_dist_matrix ~ sample_data(ps_clr)$Milk_Tier)
#No significance


#Pull out this PC1 column, combine with the GFE_Efficiency column, and then plot!
meta.rare.first <- meta[meta$parity.group == "First",]

plot_data <- plot_data[order(row.names(plot_data)),]
meta.rare.abund.2 <- meta.rare.first[order(row.names(meta.rare.first)),]

plot_data_2 <- as.data.frame(plot_data[,1:2])
row.names(plot_data_2) <- row.names(plot_data)
colnames(plot_data_2)[1] <- "PC1"
colnames(plot_data_2)[2] <- "PC2"

meta.rare.abund.3 <- merge(meta.rare.abund.2, plot_data_2, by="row.names")
row.names(meta.rare.abund.3) <- meta.rare.abund.3$Row.names


meta.rare.abund.3$DIM <- as.numeric(meta.rare.abund.3$DIM)
meta.rare.abund.3$Milk7d_T1 <- as.numeric(meta.rare.abund.3$Milk7d_T1)
meta.rare.abund.3$Darkness <- as.numeric(meta.rare.abund.3$Darkness)


##Linear modelling of GFE/RFI against PC1 axis for ALL ANIMALS
dim.pc1.1 <- ggplot(meta.rare.abund.3, aes(DIM, PC1), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("DIM x PC1") +
    theme(axis.title=element_text(size=14))

dim.pc2.1 <- ggplot(meta.rare.abund.3, aes(DIM, PC2), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("DIM x PC2") +
    theme(axis.title=element_text(size=14))




vol.pc1.1 <- ggplot(meta.rare.abund.3, aes(Milk7d_T1, PC1)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("Yield x PC1") +
    xlab("Mean Milk Volume (lbs/day)") +
    theme(axis.title=element_text(size=14))

vol.pc2.1 <- ggplot(meta.rare.abund.3, aes(Milk7d_T1, PC2)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("Yield x PC2") +
    xlab("Mean Milk Yield (lbs/day)") +
    theme(axis.title=element_text(size=14))


lm_fit_Darkness.DIM <- lm(Darkness ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_Darkness.DIM)
#0.0751 
lm_fit_Darkness.MilkVol <- lm(Darkness ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Darkness.MilkVol)
#Not significant


lm_fit_DIM.PC1 <- lm(PC1 ~ DIM + Darkness, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC1)
5#Not significant

lm_fit_Milk7d_T1.PC1 <- lm(PC1 ~ Milk7d_T1 + Darkness, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC1)
#Not significant


lm_fit_DIM.PC2 <- lm(PC2 ~ DIM + Darkness, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC2)
#SIGNIFICANT!!! p < 0.00243

lm_fit_Milk7d_T1.PC2 <- lm(PC2 ~ Milk7d_T1 + Darkness, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC2)
#SIGNIFICANT!!! p < 0.000513


lm_fit_DIM.PC1 <- lm(PC1 ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC1)
#Not significant

lm_fit_Milk7d_T1.PC1 <- lm(PC1 ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC1)
#Not significant


lm_fit_DIM.PC2 <- lm(PC2 ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC2)
#SIGNIFICANT!!! p < 0.00243

lm_fit_Milk7d_T1.PC2 <- lm(PC2 ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC2)
#SIGNIFICANT!!! p < 0.000513


grid.arrange(dim.pc1.1, dim.pc2.1, vol.pc1.1, vol.pc2.1, top = "First Lactation", ncol = 2)

ggarrange(dim.pc1.1, dim.pc2.1, vol.pc1.1, vol.pc2.1,
          ncol=2,
          nrow=2,
          labels=c("A", "B", "C", "D"),
          font.label=list(size=16))

```


##SECOND LACTATION Animals
```{r}
physeq.second <- subset_samples(physeq, PARITY == 2)

Milk7d_T1 <- sample_data(physeq.second)$Milk7d_T1
Milk7d_T1 <- Milk7d_T1[order(Milk7d_T1)]
Milk7d_T1


cutoff = round((0.2 * length(Milk7d_T1)), digits=0)

#List is sorted from lowest to highest
Milk7d_T1[cutoff]
  # 71
#Anything lower than 71 is considered Low Production

Milk7d_T1[(length(Milk7d_T1) - cutoff)]
  # 110
#Anything higher than 110 is considered High Production

sample_data(physeq.second)$Milk_Tier <- ifelse(sample_data(physeq.second)$Milk7d_T1 <= 71, "Low", ifelse(sample_data(physeq.second)$Milk7d_T1 >= 110, "High", "Mid"))


rm(list=c("Milk7d_T1", "cutoff"))

#Applying PCA to the centered log-ratio (CLR) transformed sequence counts
ps_clr <- microbiome::transform(physeq.second, "clr")

#PCA via phyloseq
ord_clr <- phyloseq::ordinate(ps_clr, "RDA")

#Plot scree plot
phyloseq::plot_scree(ord_clr) +
  geom_bar(stat="identity", fill="blue") +
  labs(x = "\nAxis", y = "Proportion of Variance\n")

#Examine eigenvalues and % proportional variance explained
head(ord_clr$CA$eig)

sapply(ord_clr$CA$eig[1:5], function(x) x / sum(ord_clr$CA$eig))     

#Scale axes and plot ordination
clr1 <- ord_clr$CA$eig[1] / sum(ord_clr$CA$eig)
clr2 <- ord_clr$CA$eig[2] / sum(ord_clr$CA$eig)
z_second <- phyloseq::plot_ordination(physeq.second, ord_clr, type="samples", color="Milk_Tier") + 
  geom_point(size = 2)  +
  coord_fixed() +  
  ylim(-12,12) +
  xlim(-12,12) +
  stat_ellipse(aes(group = Milk_Tier), linetype = 2) +
  guides(color = guide_legend(title = "Milk Volume")) +
  ggtitle("Second Lactation PCA by Milk Volume Tier") +
  scale_color_manual(values = okabe_ito_colors)

z_second

ordination_plot <- plot_ordination(physeq.second, ord_clr, type="samples", color="Milk_Tier")
plot_data <- as.data.frame(ordination_plot$data)


#Test for differences between groups
#Generate distance matrix
clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean") 
#ADONIS test
vegan::adonis2(clr_dist_matrix ~ sample_data(ps_clr)$Milk_Tier)
#No significance



#Pull out this PC1 column, combine with the GFE_Efficiency column, and then plot!
meta.rare.second <- meta[meta$parity.group == "Second",]

plot_data <- plot_data[order(row.names(plot_data)),]
meta.rare.abund.2 <- meta.rare.second[order(row.names(meta.rare.second)),]

plot_data_2 <- as.data.frame(plot_data[,1:2])
row.names(plot_data_2) <- row.names(plot_data)
colnames(plot_data_2)[1] <- "PC1"
colnames(plot_data_2)[2] <- "PC2"

meta.rare.abund.3 <- merge(meta.rare.abund.2, plot_data_2, by="row.names")
row.names(meta.rare.abund.3) <- meta.rare.abund.3$Row.names


meta.rare.abund.3$DIM <- as.numeric(meta.rare.abund.3$DIM)
meta.rare.abund.3$Milk7d_T1 <- as.numeric(meta.rare.abund.3$Milk7d_T1)
meta.rare.abund.3$Darkness <- as.numeric(meta.rare.abund.3$Darkness)


##Linear modelling of GFE/RFI against PC1 axis for ALL ANIMALS
dim.pc1.2 <- ggplot(meta.rare.abund.3, aes(DIM, PC1), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("DIM x PC1")

dim.pc2.2 <- ggplot(meta.rare.abund.3, aes(DIM, PC2), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("DIM x PC2")




vol.pc1.2 <- ggplot(meta.rare.abund.3, aes(Milk7d_T1, PC1)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("Yield x PC1") +
    xlab("Mean Milk Yield (lbs/day)")

vol.pc2.2 <- ggplot(meta.rare.abund.3, aes(Milk7d_T1, PC2)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("Yield x PC2") +
    xlab("Mean Milk Yield (lbs/day)")


lm_fit_Darkness.DIM <- lm(Darkness ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_Darkness.DIM)
#Not significant
lm_fit_Darkness.MilkVol <- lm(Darkness ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Darkness.MilkVol)
#Not significant



lm_fit_DIM.PC1 <- lm(PC1 ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC1)
#Not significant

lm_fit_Milk7d_T1.PC1 <- lm(PC1 ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC1)
#Not significant



lm_fit_DIM.PC2 <- lm(PC2 ~ DIM + Darkness, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC2)
#Not significant

lm_fit_Milk7d_T1.PC2 <- lm(PC2 ~ Milk7d_T1 + Darkness, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC2)
#Not significant

grid.arrange(dim.pc1.2, dim.pc2.2, vol.pc1.2, vol.pc2.2, top = "Second Lactation", ncol = 2)

```

##THIRD LACTATION Animals
```{r}
physeq.third <- subset_samples(physeq, PARITY == 3)

Milk7d_T1 <- sample_data(physeq.third)$Milk7d_T1
Milk7d_T1 <- Milk7d_T1[order(Milk7d_T1)]
Milk7d_T1

cutoff = round((0.2 * length(Milk7d_T1)), digits=0)

#List is sorted from lowest to highest
Milk7d_T1[cutoff]
  # 73.7
#Anything lower than 74 is considered Low Production

Milk7d_T1[(length(Milk7d_T1) - cutoff)]
  # 112
#Anything higher than 112 is considered High Production

sample_data(physeq.third)$Milk_Tier <- ifelse(sample_data(physeq.third)$Milk7d_T1 <= 74, "Low", ifelse(sample_data(physeq.third)$Milk7d_T1 >= 112, "High", "Mid"))


rm(list=c("Milk7d_T1", "cutoff"))

#Applying PCA to the centered log-ratio (CLR) transformed sequence counts
ps_clr <- microbiome::transform(physeq.third, "clr")

#PCA via phyloseq
ord_clr <- phyloseq::ordinate(ps_clr, "RDA")

#Plot scree plot
phyloseq::plot_scree(ord_clr) +
  geom_bar(stat="identity", fill="blue") +
  labs(x = "\nAxis", y = "Proportion of Variance\n")

#Examine eigenvalues and % proportional variance explained
head(ord_clr$CA$eig)

sapply(ord_clr$CA$eig[1:5], function(x) x / sum(ord_clr$CA$eig))     

#Scale axes and plot ordination
clr1 <- ord_clr$CA$eig[1] / sum(ord_clr$CA$eig)
clr2 <- ord_clr$CA$eig[2] / sum(ord_clr$CA$eig)
z_third <- phyloseq::plot_ordination(physeq.third, ord_clr, type="samples", color="Milk_Tier") + 
  geom_point(size = 2) +
  coord_fixed() +  
  ylim(-12,12) +
  xlim(-12,12) +
  stat_ellipse(aes(group = Milk_Tier), linetype = 2) +
  guides(color = guide_legend(title = "Milk Volume")) +
  ggtitle("Third Lactation PCA by Milk Volume Tier") +
  scale_color_manual(values = okabe_ito_colors)

z_third

ordination_plot <- plot_ordination(physeq.third, ord_clr, type="samples", color="Milk_Tier")
plot_data <- as.data.frame(ordination_plot$data)


#Test for differences between groups
#Generate distance matrix
clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean") 
#ADONIS test
vegan::adonis2(clr_dist_matrix ~ sample_data(ps_clr)$Milk_Tier)
#No significance



#Pull out this PC1 column, combine with the GFE_Efficiency column, and then plot!
meta.rare.third <- meta[meta$PARITY == " 3",]

plot_data <- plot_data[order(row.names(plot_data)),]
meta.rare.abund.2 <- meta.rare.third[order(row.names(meta.rare.third)),]

plot_data_2 <- as.data.frame(plot_data[,1:2])
row.names(plot_data_2) <- row.names(plot_data)
colnames(plot_data_2)[1] <- "PC1"
colnames(plot_data_2)[2] <- "PC2"

meta.rare.abund.3 <- merge(meta.rare.abund.2, plot_data_2, by="row.names")
row.names(meta.rare.abund.3) <- meta.rare.abund.3$Row.names


meta.rare.abund.3$DIM <- as.numeric(meta.rare.abund.3$DIM)
meta.rare.abund.3$Milk7d_T1 <- as.numeric(meta.rare.abund.3$Milk7d_T1)
meta.rare.abund.3$Darkness <- as.numeric(meta.rare.abund.3$Darkness)


##Linear modelling of GFE/RFI against PC1 axis for ALL ANIMALS
dim.pc1.3 <- ggplot(meta.rare.abund.3, aes(DIM, PC1), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("DIM x PC1")

dim.pc2.3 <- ggplot(meta.rare.abund.3, aes(DIM, PC2), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("DIM x PC2")




vol.pc1.3 <- ggplot(meta.rare.abund.3, aes(Milk7d_T1, PC1)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("Yield x PC1") +
    xlab("Mean Milk Yield (lbs/day)")

vol.pc2.3 <- ggplot(meta.rare.abund.3, aes(Milk7d_T1, PC2)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("Yield x PC2") +
    xlab("Mean Milk Yield (lbs/day)")



lm_fit_Darkness.DIM <- lm(Darkness ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_Darkness.DIM)
#Not significant
lm_fit_Darkness.MilkVol <- lm(Darkness ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Darkness.MilkVol)
#Not significant



lm_fit_DIM.PC1 <- lm(PC1 ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC1)
#Not significant

lm_fit_Milk7d_T1.PC1 <- lm(PC1 ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC1)
#Not significant



lm_fit_DIM.PC2 <- lm(PC2 ~ DIM, data=meta.rare.abund.3)
summary(lm_fit_DIM.PC2)
#Not significant

lm_fit_Milk7d_T1.PC2 <- lm(PC2 ~ Milk7d_T1, data=meta.rare.abund.3)
summary(lm_fit_Milk7d_T1.PC2)
#Not significant

grid.arrange(dim.pc1.3, dim.pc2.3, vol.pc1.3, vol.pc2.3, top = "Third Lactation", ncol = 2)

grid.arrange(z_first, z_second, z_third, ncol=3)

ggarrange(z_first, z_second, z_third,
          nrow=1,
          ncol=3,
          labels=c("A", "B", "C"))

```

#Alpha Diveristy Analysis
##Shannon
```{r}
physeq.swab1.abund <- subset_samples(physeq.swab1.abund, PARITY <= 3)

sample_data(physeq.swab1.abund)$Milk_Tier <- factor(sample_data(physeq.swab1.abund)$Milk_Tier, levels = c("Low", "Mid", "High"))
sample_data(physeq.swab1.abund)$Parity.Stage <- factor(sample_data(physeq.swab1.abund)$Parity.Stage, levels = c("First.Early", "First.Middle", "First.Late", "Second.Early", "Second.Middle", "Second.Late", "Third.Early", "Third.Middle", "Third.Late"))
sample_data(physeq.swab1.abund)$STAGE <- factor(sample_data(physeq.swab1.abund)$STAGE, levels = c("Early", "Middle", "Late"))
sample_data(physeq.swab1.abund)$parity.group <- factor(sample_data(physeq.swab1.abund)$parity.group, levels = c("First", "Second", "Third"))


#None of the metrics are normally distributed. Using Kruskal-Wallis test, then testing pairwise with Wilcoxon Rank Sum test.
kruskal.test(Shannon ~ Milk_Tier, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Shannon, sample_data(physeq.swab1.abund)$Milk_Tier, p.adjust.method="fdr")
#      Low  Mid 
# Mid  0.58 -   
# High 0.58 0.85

b_p1 <- ggplot(sample_data(physeq.swab1.abund), aes(x=Milk_Tier, y=Shannon)) + 
  geom_boxplot() +
  labs(title=,x="Swab Milk Tiers", y = " Bacterial Shannon Diversity") + 
  theme(axis.text=element_text(size=12)) +  
  ylim(2.25, 7.5)
b_p1


#parity.stage
kruskal.test(Shannon ~ Parity.Stage, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Shannon, sample_data(physeq.swab1.abund)$Parity.Stage, p.adjust.method="fdr")
#               First.Early First.Middle First.Late Second.Early Second.Middle Second.Late Third.Early Third.Middle
# First.Middle  1.00        -            -          -            -             -           -           -           
# First.Late    0.74        0.85         -          -            -             -           -           -           
# Second.Early  0.23        0.74         0.74       -            -             -           -           -           
# Second.Middle 0.74        0.74         0.74       0.23         -             -           -           -           
# Second.Late   0.74        1.00         1.00       0.74         0.74          -           -           -           
# Third.Early   0.83        0.79         0.74       0.23         1.00          0.74        -           -           
# Third.Middle  0.79        0.89         0.74       0.23         0.83          0.74        1.00        -           
# Third.Late    0.74        0.74         0.74       0.23         0.85          0.74        1.00        1.00 

b_p2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=Parity.Stage, y=Shannon)) + 
  geom_boxplot() +
  labs(x="", y = "Shannon Diversity") + 
  theme(axis.text = element_text(size=17)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(axis.title = element_text(size=20))
  
b_p2



#Stage
kruskal.test(Shannon ~ STAGE, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Shannon, sample_data(physeq.swab1.abund)$STAGE, p.adjust.method="fdr")
#        Early Middle
# Middle 0.94  -     
# Late   0.94  0.94  

b_p3 <- ggplot(sample_data(physeq.swab1.abund), aes(x=STAGE, y=Shannon)) + 
  geom_boxplot() +
  labs(title=,x="Swab Stages", y = "Bacterial Shannon Diversity") + 
  theme(axis.text=element_text(size=12))
b_p3


#Lactation
kruskal.test(Shannon ~ parity.group, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Shannon, sample_data(physeq.swab1.abund)$parity.group, p.adjust.method="fdr")
#        First Second
# Second 0.98  -     
# Multi  0.98  0.98


b_p4 <- ggplot(sample_data(physeq.swab1.abund), aes(x=parity.group, y=Shannon)) + 
  geom_boxplot() +
  labs(title=,x="Swab Lactations", y = "Bacterial Shannon Diversity") + 
  theme(axis.text=element_text(size=12))
b_p4

```

##Correlating DIM and Milk Production to Shannon Scores
```{r}
physeq.swab1 <- subset_samples(physeq.swab1.abund, Swab_Rumen=="Swab")
sample_data(physeq.swab1)$DIM <- as.numeric(sample_data(physeq.swab1)$DIM)
sample_data(physeq.swab1)$Milk7d_T1 <- as.numeric(sample_data(physeq.swab1)$Milk7d_T1)


b_p5 <- ggplot(sample_data(physeq.swab1), aes(x = DIM, y = Shannon)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Days in Milk", y = "Bacterial Shannon Diversity") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_p5

b_p6 <- ggplot(sample_data(physeq.swab1), aes(x = Milk7d_T1, y = Shannon)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Days in Milk", y = "Bacterial Shannon Diversity") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_p6


#Days in Milk
cor.test(sample_data(physeq.swab1)$DIM, sample_data(physeq.swab1)$Shannon, method = "spearman", exact = FALSE)
# data:  sample_data(physeq.swab1)$DIM and sample_data(physeq.swab1)$Shannon
# S = 5069954, p-value = 0.5243
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.0359527

cor.test(sample_data(physeq.swab1)$DIM, sample_data(physeq.swab1)$Shannon, method = "kendall")
# data:  sample_data(physeq.swab1)$DIM and sample_data(physeq.swab1)$Shannon
# z = 0.68893, p-value = 0.4909
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#        tau 
# 0.02601743 


#Milk Production
cor.test(sample_data(physeq.swab1)$Milk7d_T1, sample_data(physeq.swab1)$Shannon, method = "spearman", exact = FALSE)
# data:  sample_data(physeq.swab1)$Milk7d_T1 and sample_data(physeq.swab1)$Shannon
# S = 5208373, p-value = 0.8646
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#         rho 
# 0.009632298

cor.test(sample_data(physeq.swab1)$Milk7d_T1, sample_data(physeq.swab1)$Shannon, method = "kendall")
# data:  sample_data(physeq.swab1)$Milk7d_T1 and sample_data(physeq.swab1)$Shannon
# z = 0.17104, p-value = 0.8642
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#         tau 
# 0.006461819

```


##Inverse Simpson
```{r}
physeq.swab1.abund <- subset_samples(physeq.swab1.abund, PARITY <= 3)

sample_data(physeq.swab1.abund)$Milk_Tier <- factor(sample_data(physeq.swab1.abund)$Milk_Tier, levels = c("Low", "Mid", "High"))
sample_data(physeq.swab1.abund)$Parity.Stage <- factor(sample_data(physeq.swab1.abund)$Parity.Stage, levels = c("First.Early", "First.Middle", "First.Late", "Second.Early", "Second.Middle", "Second.Late", "Third.Early", "Third.Middle", "Third.Late"))
sample_data(physeq.swab1.abund)$STAGE <- factor(sample_data(physeq.swab1.abund)$STAGE, levels = c("Early", "Middle", "Late"))
sample_data(physeq.swab1.abund)$parity.group <- factor(sample_data(physeq.swab1.abund)$parity.group, levels = c("First", "Second", "Third"))


#None of the metrics are normally distributed. Using Kruskal-Wallis test, then testing pairwise with Wilcoxon Rank Sum test.
kruskal.test(InvSimpson ~ Milk_Tier, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$InvSimpson, sample_data(physeq.swab1.abund)$Milk_Tier, p.adjust.method="fdr")
#      Low  Mid 
# Mid  0.87 -   
# High 0.87 0.87

b_p7 <- ggplot(sample_data(physeq.swab1.abund), aes(x=Milk_Tier, y=InvSimpson)) + 
  geom_boxplot() +
  labs(title=,x="Swab Milk Tiers", y = " Bacterial InvSimpson Diversity") + 
  theme(axis.text=element_text(size=12))
b_p7


#parity.stage
kruskal.test(InvSimpson ~ Parity.Stage, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$InvSimpson, sample_data(physeq.swab1.abund)$Parity.Stage, p.adjust.method="fdr")
#               First.Early First.Middle First.Late Second.Early Second.Middle Second.Late Multi.Early Multi.Middle
# First.Middle  0.97        -            -          -            -             -           -           -           
# First.Late    0.95        0.97         -          -            -             -           -           -           
# Second.Early  0.60        0.84         0.93       -            -             -           -           -           
# Second.Middle 0.67        0.84         0.60       0.40         -             -           -           -           
# Second.Late   0.97        0.97         0.97       0.62         0.60          -           -           -           
# Multi.Early   0.60        0.93         0.95       0.97         0.40          0.67        -           -           
# Multi.Middle  0.97        0.97         0.97       0.60         0.60          0.97        0.60        -           
# Multi.Late    0.97        0.97         0.95       0.60         0.60          0.97        0.62        0.97

b_p8 <- ggplot(sample_data(physeq.swab1.abund), aes(x=Parity.Stage, y=InvSimpson)) + 
  geom_boxplot() +
  labs(title=,x="", y = "Inverse Simpson") + 
  theme(axis.text = element_text(size=17)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(axis.title = element_text(size=20))
b_p8



#Stage
kruskal.test(InvSimpson ~ STAGE, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$InvSimpson, sample_data(physeq.swab1.abund)$STAGE, p.adjust.method="fdr")
#        Early Middle
# Middle 0.20  -     
# Late   0.46  0.46  

b_p9 <- ggplot(sample_data(physeq.swab1.abund), aes(x=STAGE, y=InvSimpson)) + 
  geom_boxplot() +
  labs(title=,x="Swab Stages", y = "Bacterial InvSimpson Diversity") + 
  theme(axis.text=element_text(size=12))
b_p9


#Lactation
kruskal.test(InvSimpson ~ parity.group, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$InvSimpson, sample_data(physeq.swab1.abund)$parity.group, p.adjust.method="fdr")
#        First Second
# Second 0.98  -     
# Multi  0.98  0.98


b_p10 <- ggplot(sample_data(physeq.swab1.abund), aes(x=parity.group, y=InvSimpson)) + 
  geom_boxplot() +
  labs(title=,x="Swab Lactations", y = "Bacterial InvSimpson Diversity") + 
  theme(axis.text=element_text(size=12))
b_p10

```

##Correlating DIM and Milk Production to Inverse Simpson
```{r}
physeq.swab1 <- subset_samples(physeq.swab1.abund, Swab_Rumen=="Swab")
sample_data(physeq.swab1)$DIM <- as.numeric(sample_data(physeq.swab1)$DIM)
sample_data(physeq.swab1)$Milk7d_T1 <- as.numeric(sample_data(physeq.swab1)$Milk7d_T1)


b_p11 <- ggplot(sample_data(physeq.swab1), aes(x = DIM, y = InvSimpson)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Days in Milk", y = "Bacterial InvSimpson Diversity") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_p11

b_p12 <- ggplot(sample_data(physeq.swab1), aes(x = Milk7d_T1, y = InvSimpson)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Days in Milk", y = "Bacterial InvSimpson Diversity") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_p12


#Days in Milk
cor.test(sample_data(physeq.swab1)$DIM, sample_data(physeq.swab1)$InvSimpson, method = "spearman", exact = FALSE)
# data:  sample_data(physeq.swab1)$DIM and sample_data(physeq.swab1)$InvSimpson
# S = 4978202, p-value = 0.3441
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#        rho 
# 0.05339919

cor.test(sample_data(physeq.swab1)$DIM, sample_data(physeq.swab1)$InvSimpson, method = "kendall")
# data:  sample_data(physeq.swab1)$DIM and sample_data(physeq.swab1)$InvSimpson
# z = 1.0065, p-value = 0.3142
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#     tau 
# 0.03801

#Milk Production
cor.test(sample_data(physeq.swab1)$Milk7d_T1, sample_data(physeq.swab1)$InvSimpson, method = "spearman", exact = FALSE)
# data:  sample_data(physeq.swab1)$Milk7d_T1 and sample_data(physeq.swab1)$InvSimpson
# S = 5454379, p-value = 0.5106
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#         rho 
# -0.03714544 

cor.test(sample_data(physeq.swab1)$Milk7d_T1, sample_data(physeq.swab1)$InvSimpson, method = "kendall")
# data:  sample_data(physeq.swab1)$Milk7d_T1 and sample_data(physeq.swab1)$InvSimpson
# z = -0.67616, p-value = 0.4989
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#         tau 
# -0.02554532 

```


##Chao
```{r}
physeq.swab1.abund <- subset_samples(physeq.swab1.abund, PARITY <= 3)

sample_data(physeq.swab1.abund)$Milk_Tier <- factor(sample_data(physeq.swab1.abund)$Milk_Tier, levels = c("Low", "Mid", "High"))
sample_data(physeq.swab1.abund)$Parity.Stage <- factor(sample_data(physeq.swab1.abund)$Parity.Stage, levels = c("First.Early", "First.Middle", "First.Late", "Second.Early", "Second.Middle", "Second.Late", "Third.Early", "Third.Middle", "Third.Late"))
sample_data(physeq.swab1.abund)$STAGE <- factor(sample_data(physeq.swab1.abund)$STAGE, levels = c("Early", "Middle", "Late"))
sample_data(physeq.swab1.abund)$parity.group <- factor(sample_data(physeq.swab1.abund)$parity.group, levels = c("First", "Second", "Third"))


#None of the metrics are normally distributed. Using Kruskal-Wallis test, then testing pairwise with Wilcoxon Rank Sum test.
kruskal.test(InvSimpson ~ Chao1, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Chao1, sample_data(physeq.swab1.abund)$Milk_Tier, p.adjust.method="fdr")
#      Low  Mid 
# Mid  0.87 -   
# High 0.87 0.87

b_p13 <- ggplot(sample_data(physeq.swab1.abund), aes(x=Milk_Tier, y=Chao1)) + 
  geom_boxplot() +
  labs(title=,x="Swab Milk Tiers", y = " Bacterial Chao1 Diversity") + 
  theme(axis.text=element_text(size=12))
b_p13


#parity.stage
kruskal.test(Chao1 ~ Parity.Stage, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Chao1, sample_data(physeq.swab1.abund)$Parity.Stage, p.adjust.method="fdr")
#               First.Early First.Middle First.Late Second.Early Second.Middle Second.Late Multi.Early Multi.Middle
# First.Middle  0.90        -            -          -            -             -           -           -           
# First.Late    0.90        0.90         -          -            -             -           -           -           
# Second.Early  0.87        0.88         0.87       -            -             -           -           -           
# Second.Middle 0.87        0.87         0.90       0.59         -             -           -           -           
# Second.Late   0.99        0.90         0.90       0.87         0.87          -           -           -           
# Multi.Early   0.93        0.90         0.90       0.87         0.87          0.96        -           -           
# Multi.Middle  0.87        0.87         0.93       0.59         0.90          0.87        0.87        -           
# Multi.Late    0.90        0.87         0.90       0.59         0.90          0.87        0.87        0.99

b_p14 <- ggplot(sample_data(physeq.swab1.abund), aes(x=Parity.Stage, y=Chao1)) + 
  geom_boxplot() +
  labs(x="", y = "Chao Richness") + 
  theme(axis.text = element_text(size=17)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(axis.title = element_text(size=20))
b_p14



#Stage
kruskal.test(Chao1 ~ STAGE, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Chao1, sample_data(physeq.swab1.abund)$STAGE, p.adjust.method="fdr")
#        Early Middle
# Middle 0.24  -     
# Late   0.24  0.79  

b_p15 <- ggplot(sample_data(physeq.swab1.abund), aes(x=STAGE, y=Chao1)) + 
  geom_boxplot() +
  labs(title=,x="Swab Stages", y = "Bacterial Chao1 Diversity") + 
  theme(axis.text=element_text(size=12))
b_p15


#Lactation
kruskal.test(Chao1 ~ parity.group, data=as.matrix(sample_data(physeq.swab1.abund)))
# pvalue > 0.05, NO sig difference
pairwise.wilcox.test(sample_data(physeq.swab1.abund)$Chao1, sample_data(physeq.swab1.abund)$parity.group, p.adjust.method="fdr")
#        First Second
# Second 0.98  -     
# Multi  0.98  0.98


b_p16 <- ggplot(sample_data(physeq.swab1.abund), aes(x=parity.group, y=Chao1)) + 
  geom_boxplot() +
  labs(title=,x="Swab Lactations", y = "Bacterial Chao1 Diversity") + 
  theme(axis.text=element_text(size=12))
b_p16

```

##Correlating DIM and Milk Production to Chao
```{r}
physeq.swab1 <- subset_samples(physeq.swab1.abund, Swab_Rumen=="Swab")
sample_data(physeq.swab1)$DIM <- as.numeric(sample_data(physeq.swab1)$DIM)
sample_data(physeq.swab1)$Milk7d_T1 <- as.numeric(sample_data(physeq.swab1)$Milk7d_T1)


b_p17 <- ggplot(sample_data(physeq.swab1), aes(x = DIM, y = Chao1)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Days in Milk", y = "Bacterial Chao1 Diversity") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_p17

b_p18 <- ggplot(sample_data(physeq.swab1), aes(x = Milk7d_T1, y = Chao1)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Days in Milk", y = "Bacterial Chao1 Diversity") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_p18


#Days in Milk
cor.test(sample_data(physeq.swab1)$DIM, sample_data(physeq.swab1)$Chao1, method = "spearman", exact = FALSE)
# data:  sample_data(physeq.swab1)$DIM and sample_data(physeq.swab1)$Chao1
# S = 4935546, p-value = 0.2757
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#        rho 
# 0.06151015 

cor.test(sample_data(physeq.swab1)$DIM, sample_data(physeq.swab1)$Chao1, method = "kendall")
# data:  sample_data(physeq.swab1)$DIM and sample_data(physeq.swab1)$Chao1
# z = 1.1104, p-value = 0.2668
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#        tau 
# 0.04193416

#Milk Production
cor.test(sample_data(physeq.swab1)$Milk7d_T1, sample_data(physeq.swab1)$Chao1, method = "spearman", exact = FALSE)
# data:  sample_data(physeq.swab1)$Milk7d_T1 and sample_data(physeq.swab1)$Chao1
# S = 5057490, p-value = 0.4973
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#        rho 
# 0.03832269 

cor.test(sample_data(physeq.swab1)$Milk7d_T1, sample_data(physeq.swab1)$Chao1, method = "kendall")
# data:  sample_data(physeq.swab1)$Milk7d_T1 and sample_data(physeq.swab1)$Chao1
# z = 0.75874, p-value = 0.448
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.0286658

```

#Creating plot that has parity.stage comparisons for alpha diversities
```{r}
grid.arrange(b_p2, b_p8, b_p14, ncol = 3)

ggarrange(b_p2, b_p8, b_p14,
          nrow=1,
          ncol=3,
          labels=c("A", "B", "C"))

```

















