---
title: "Second Lactation Alpha Diversity, Linear Modelling, and DESeq2 Analysis of GFE and RFI"
author: "Joseph Skarlupka V"
date: "2024-10-31"
output: html_document
---

#Loading Libraries
```{r, Libraries}
library("ggplot2")
library("phyloseq")
library("vegan")
library("DESeq2")
library("reltools")
library("phylosmith")

set.seed(81471438)
```

#Importing files of interest
```{r, import the files of interest}
#Load in the phyloseq object that contains
  #Time point 1
  #Swabs only
  #Only with MPE information
  #7000 sequences
  #2nd Lactation Arlington Animals

load("Second Lactation RFI and GFE Animals.RData")

```

#Changing values to integers as needed
```{r}
#A lot of the work I did ended up changing the class of some rows from integers to characters. Making the necessary changes back to integers here

sample_data(physeq)$rfi_recent <- as.numeric(sample_data(physeq)$rfi_recent)


```


#RFI - Overwrite the rfi_Efficiency column to reflect the top and bottom 20% animals of only the first lactation
```{r}
RFI_List <- sample_data(physeq)$rfi_recent
RFI_List <- RFI_List[order(RFI_List)]
RFI_List

cutoff = round((0.2 * length(RFI_List)), digits=0)

#List is sorted from lowest to highest
RFI_List[cutoff]
  # -1.9
#Anything lower than -1.9 is considered HE

RFI_List[(length(RFI_List) - cutoff)]
  # 0.9
#Anything higher than 0.9 is considered LE

sample_data(physeq)$rfi_Efficiency <- ifelse(sample_data(physeq)$rfi_recent <= -1.9, "HE", ifelse(sample_data(physeq)$rfi_recent >= 0.9, "LE", "MID"))

sample_data(physeq)$Sample.Type_RFI.Efficiency <- paste0(sample_data(physeq)$Sample.Type, "_", sample_data(physeq)$rfi_Efficiency)
table(sample_data(physeq)$Sample.Type_RFI.Efficiency)

rm(list=c("RFI_List", "cutoff"))


```



#GFE - Split in top and bottom 20%
```{r, Split Animals into 20% HE and LE groups for GFE}
#Applying HE and LE values to meta based on GFE

# Take the highest 20% of GFEs, and classify those as high efficiency
# Take the lowest 20% of GFEs, and classify those as low efficiency
# Classify the rest as mid efficiency

GFEs <- data.frame(sample_data(physeq)$Cow, sample_data(physeq)$GFE)

GFE_List <- as.numeric(GFEs$sample_data.physeq..GFE)
GFE_List <- GFE_List[order(GFE_List)]
GFE_List

cutoff = round((0.2 * length(GFE_List)), digits=0)

#List is sorted from lowest to highest
GFE_List[cutoff]
  # 1.0714286 
#Anything lower than 1.0714286  is considered LE

GFE_List[(length(GFE_List) - cutoff)]
  # 1.2000000
#Anything higher than 1.2000000 is considered HE

sample_data(physeq)$gfe_20percent <- ifelse(sample_data(physeq)$GFE <= 1.0714286 , "LE", ifelse(sample_data(physeq)$GFE >= 1.2000000, "HE", "MID"))

sample_data(physeq)$Sample.Type_GFE.Efficiency <- paste0(sample_data(physeq)$Sample.Type, "_", sample_data(physeq)$gfe_20percent)

table(sample_data(physeq)$Sample.Type_GFE.Efficiency)
 # Swab_HE  Swab_LE Swab_MID 
 #       6        6       12

rm(list=c("GFEs", "GFE_List", "cutoff"))


```

#Rarefying the phyloseq object
```{r}
rowSums(otu_table(physeq))

#Rarefying
physeq.swab1.rare <- rarefy_even_depth(physeq, sample.size=7000)
table(sample_data(physeq.swab1.rare)$parity.group)
table(sample_data(physeq.swab1.rare)$Sample.Type_GFE.Efficiency)
table(sample_data(physeq.swab1.rare)$Sample.Type_RFI.Efficiency)

#7000 - 30 samples removed

tab <- otu_table(physeq.swab1.rare)
class(tab) <- "matrix"
rarecurve(tab, step=50, cex=0.5, xlim=c(0,15000), ylim=c(0,2000))
rm(tab)

#Pulling back out the rarefied ASV table with new column names
ASV1 = as(otu_table(physeq.swab1.rare), "matrix")
if(taxa_are_rows(physeq.swab1.rare)){ASV1 <- t(ASV1)}
# Coerce to data.frame
ASV.7000 = as.data.frame(ASV1)
rm(ASV1)


#Pulling back out the metadata table with new column names
meta.7000 = as(sample_data(physeq.swab1.rare), "matrix")
# Coerce to data.frame
meta.7000 = as.data.frame(meta.7000)


#Pulling out new taxonomy file
tax.7000 = as(tax_table(physeq.swab1.rare), "matrix")
tax.7000 = as.data.frame(tax.7000)

```


##Abundance Cutoff of Phyloseq Object
```{r}
#Removing ASVs from which there are no samples with more than 10 sequences
physeq.swab1.abund <- prune_taxa(taxa_sums(physeq.swab1.rare)>10,physeq.swab1.rare)


```

##Pulling out the ASV, Meta, and Taxonomy of the rarefied and abundance cutoff phyloseq object
```{r}
#Pulling back out the rarefied ASV table with new column names
ASV1 = as(otu_table(physeq.swab1.abund), "matrix")
if(taxa_are_rows(physeq.swab1.abund)){ASV1 <- t(ASV1)}
# Coerce to data.frame
ASV.rare.abund = as.data.frame(ASV1)
rm(ASV1)


#Pulling back out the metadata table with new column names
meta.rare.abund = as(sample_data(physeq.swab1.abund), "matrix")
# Coerce to data.frame
meta.rare.abund = as.data.frame(meta.rare.abund)


#Pulling out new taxonomy file
tax.rare.abund = as(tax_table(physeq.swab1.abund), "matrix")
tax.rare.abund = as.data.frame(tax.rare.abund)

```


##Creating a phyloseq object of that includes an abundance cutoff, but no rarefaction
```{r}
#Removing ASVs from which there are no samples with more than 10 sequences

physeq.norare.abund <- prune_taxa(taxa_sums(physeq)>10,physeq)


```


##Adding alpha diversity metrics in phyloseq
```{r, adding alpha diversity metrics in phyloseq, then pulling out the component tables again so I can do some work in vegan}
#putting all metrics into a file
div <- phyloseq::estimate_richness(physeq.swab1.abund, measures=c("Observed", "Chao1", "ACE", "Shannon", "InvSimpson", "Simpson"))
row.names(div) <- gsub(".", "-", row.names(div), fixed=TRUE)
physeq.swab1.abund <- merge_phyloseq(physeq.swab1.abund, sample_data(div))

rm(list=c("div"))

```


#Exploring Alpha Diversity for GFE and RFI
```{r, Exploring alpha diversity of the efficiency groups when split into top and bottom 20% by GFE}
plot_richness(physeq.swab1.abund, "gfe_20percent", measures="Shannon")

#Testing normality
shapiro.test(sample_data(physeq.swab1.abund)$Shannon)
#W = 0.96688, p-value = 0.5908
shapiro.test(sample_data(physeq.swab1.abund)$Chao1)
#W = 0.96126, p-value = 0.4642
shapiro.test(sample_data(physeq.swab1.abund)$ACE)
#W = 0.96249, p-value = 0.4905
shapiro.test(sample_data(physeq.swab1.abund)$Simpson)
#W = 0.90973, p-value = 0.0348
shapiro.test(sample_data(physeq.swab1.abund)$InvSimpson)
#W = 0.84416, p-value = 0.001707
shapiro.test(sample_data(physeq.swab1.abund)$Observed)
#W = 0.95515, p-value = 0.3489

par(mfrow=c(2,3))


#None of the metrics are normally distributed. Using Kruskal-Wallis test, then testing pairwise with Wilcoxon Rank Sum test.
#The RFI and GFE labels of HE/LE/Mid are categorical
##GFE

#SHANNON
aov.shannon.gfe <- aov(Shannon ~ gfe_20percent, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.shannon.gfe)
#               Df Sum Sq Mean Sq F value Pr(>F)  
# gfe_20percent  2  4.609  2.3045   5.629  0.011 *
# Residuals     21  8.597  0.4094
TukeyHSD(aov.shannon.gfe)
# $gfe_20percent
#               diff        lwr         upr     p adj
# LE-HE  -0.94268333 -1.8738149 -0.01155172 0.0468818
# MID-HE  0.09773617 -0.7086475  0.90411980 0.9499753
# MID-LE  1.04041950  0.2340359  1.84680313 0.0102100

    #DOES TUBE COLOR AFFECT THE MODEL?
aov.shannon.gfe.tube = aov(Shannon ~ gfe_20percent*Tube.Color, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.shannon.gfe.tube)
  #                          Df Sum Sq Mean Slq F value Pr(>F)  
  # gfe_20percent             2  4.362  2.1809   5.778 0.0148 *
  # Tube.Color                3  1.502  0.5007   1.327 0.3053  
  # gfe_20percent:Tube.Color  3  1.056  0.3520   0.933 0.4509  
  # Residuals                14  5.284  0.3774
    #NO IT DOESN'T
aov.shannon.gfe.tube = aov(Shannon ~ gfe_20percent+Tube.Color, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.shannon.gfe.tube)
  #               Df Sum Sq Mean Sq F value Pr(>F)  
  # gfe_20percent  2  4.362  2.1809   5.848 0.0117 *
  # Tube.Color     3  1.502  0.5007   1.343 0.2937  
  # Residuals     17  6.340  0.3729                 

#CHAO
aov.chao.gfe <- aov(Chao1 ~ gfe_20percent, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.chao.gfe)
# p >0.05, no difference
TukeyHSD(aov.chao.gfe)


#ACE
aov.ACE.gfe <- aov(ACE ~ gfe_20percent, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.ACE.gfe)
# p >0.05, no difference

#SIMPSON
kruskal.test(Simpson ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)))
# p >0.05, no difference

#INVSIMPSON
kruskal.test(InvSimpson ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)))
# p >0.05, no difference

#OBSERVED
aov.observed.gfe <- aov(Observed ~ gfe_20percent, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.observed.gfe)
#               Df Sum Sq Mean Sq F value Pr(>F)  
# gfe_20percent  2 180099   90049    3.81 0.0388 *
# Residuals     21 496376   23637  
TukeyHSD(aov.observed.gfe)
# $gfe_20percent
#              diff         lwr       upr     p adj
# LE-HE  -192.33333 -416.068403  31.40174 0.1005951
# MID-HE   11.16667 -182.593587 204.92692 0.9884404
# MID-LE  203.50000    9.739746 397.26025 0.0384291
    
    #DOES TUBE COLOR AFFECT THE MODEL?
aov.observed.gfe.tube = aov(Observed ~ gfe_20percent*Tube.Color, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.observed.gfe.tube)
    #                          Df Sum Sq Mean Sq F value Pr(>F)  
    # gfe_20percent             2 165680   82840   3.992 0.0425 *
    # Tube.Color                3  64960   21653   1.043 0.4039  
    # gfe_20percent:Tube.Color  3  53834   17945   0.865 0.4823  
    # Residuals                14 290534   20752    

#NO IT DOESN'T

aov.observed.gfe.tube = aov(Observed ~ gfe_20percent+Tube.Color, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.observed.gfe.tube)
    #               Df Sum Sq Mean Sq F value Pr(>F)  
    # gfe_20percent  2 165680   82840   4.089 0.0355 *
    # Tube.Color     3  64960   21653   1.069 0.3884  
    # Residuals     17 344367   20257                 


boxplot(as.numeric(Shannon) ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Shannon diversity")
boxplot(as.numeric(Chao1) ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Chao1 Richness")
boxplot(as.numeric(ACE) ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="ACE")
boxplot(as.numeric(Simpson) ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Simpson")
boxplot(as.numeric(InvSimpson) ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="InvSimpson")
boxplot(as.numeric(Observed) ~ gfe_20percent, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Observed")

p1.2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=gfe_20percent, y=Shannon)) + 
  geom_boxplot() +
  labs(x="GFE Groups", y = "Shannon Diversity") +
  scale_x_discrete(limits = c("LE", "HE", "MID"))
p1.2

p2.2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=gfe_20percent, y=Chao1)) + 
  geom_boxplot() +
  labs(x="GFE Groups", y = "Chao Richness") +
  scale_x_discrete(limits = c("LE", "HE", "MID")) 
p2.2

p3.2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=gfe_20percent, y=InvSimpson)) + 
  geom_boxplot() +
  labs(x="GFE Groups", y = "Inverse Simpson") +
  scale_x_discrete(limits = c("LE", "HE", "MID")) 
p3.2


rfi_Efficiency
##RFI
aov.shannon.rfi <- aov(Shannon ~ rfi_Efficiency, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.shannon.rfi)
# p = 0.0461, Significant difference
TukeyHSD(aov.shannon.rfi)
    #               diff        lwr         upr     p adj
    # LE-HE   0.93644740 -0.1554715  2.02836627 0.1015737
    # MID-HE  0.02196899 -0.8775041  0.92144204 0.9979129
    # MID-LE -0.91447841 -1.8139515 -0.01500536 0.0458486
aov.shannon.rfi.tube = aov(Shannon ~ rfi_Efficiency*Tube.Color, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.shannon.rfi.tube)
    #                           Df Sum Sq Mean Sq F value Pr(>F)  
    # rfi_Efficiency             2  4.116  2.0580   5.270 0.0197 *
    # Tube.Color                 3  1.262  0.4208   1.078 0.3905  
    # rfi_Efficiency:Tube.Color  3  1.359  0.4529   1.160 0.3599  
    # Residuals                 14  5.467  0.3905    

#NO IT DOESN'T

aov.shannon.rfi.tube = aov(Shannon ~ rfi_Efficiency+Tube.Color, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.shannon.rfi.tube)
  #                Df Sum Sq Mean Sq F value Pr(>F)  
  # rfi_Efficiency  2  3.334  1.6669   3.523 0.0512 .
  # Tube.Color      3  1.327  0.4422   0.934 0.4445  
  # Residuals      18  8.518  0.4732                 


aov.chao.rfi <- aov(Chao1 ~ rfi_Efficiency, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.chao.rfi)
# p > 0.05, No difference
aov.ACE.rfi <- aov(ACE ~ rfi_Efficiency, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.ACE.rfi)
# p >0.05, no difference
kruskal.test(Simpson ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)))
# p >0.05, no difference
kruskal.test(InvSimpson ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)))
# p = 0.05698 
kruskal.test(InvSimpson ~ Tube.Color, data=as.matrix(sample_data(physeq.swab1.abund)))

aov.observed.rfi <- aov(Observed ~ rfi_Efficiency, data=as.data.frame(as.matrix(sample_data(physeq.swab1.abund))))
summary(aov.observed.rfi)
# p >0.05, no difference

boxplot(as.numeric(Shannon) ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Shannon diversity")
boxplot(as.numeric(Chao1) ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Chao1 Richness")
boxplot(as.numeric(ACE) ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="ACE")
boxplot(as.numeric(Simpson) ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Simpson")
boxplot(as.numeric(InvSimpson) ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="InvSimpson")
boxplot(as.numeric(Observed) ~ rfi_Efficiency, data=as.matrix(sample_data(physeq.swab1.abund)), ylab="Observed")

p4.2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=rfi_Efficiency, y=Shannon)) + 
  geom_boxplot() +
  labs(x="RFI Groups", y = "Shannon Diversity") +
  scale_x_discrete(limits = c("LE", "HE", "MID"))
p4.2

p5.2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=rfi_Efficiency, y=Chao1)) + 
  geom_boxplot() +
  labs(x="RFI Groups", y = "Chao Richness") +
  scale_x_discrete(limits = c("LE", "HE", "MID")) 
p5.2

p6.2 <- ggplot(sample_data(physeq.swab1.abund), aes(x=rfi_Efficiency, y=InvSimpson)) + 
  geom_boxplot() +
  labs(x="RFI Groups", y = "Inverse Simpson") +
  scale_x_discrete(limits = c("LE", "HE", "MID")) 
p6.2

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3, top="Alpha Diversities of Second Lactation")


```


#Pulling out the ASV table from the phyloseq object that has had abundance cutoff but has NOT been rarefied
```{r}
#Pulling out the rarefied ASV table fropm the phyloseq object that has been rarefied and has had abundance cutoff done
  #This is the same dataset that the DESeq2 work was done on

ASV.norare.abund = as(otu_table(physeq.norare.abund), "matrix")
if(taxa_are_rows(physeq.norare.abund)){ASV.norare.abund <- t(ASV.norare.abund)}
# Coerce to data.frame
ASV.norare.abund = as.data.frame(ASV.norare.abund)

#Pulling back out the metadata table with new column names
meta.norare.abund = as(sample_data(physeq.norare.abund), "matrix")
# Coerce to data.frame
meta.norare.abund = as.data.frame(meta.norare.abund)


#Pulling out new taxonomy file
tax.norare.abund = as(tax_table(physeq.norare.abund), "matrix")
tax.norare.abund = as.data.frame(tax.norare.abund)

```

#Testing if distribution of numbers for tube colors is different between high and low efficiency animals.
```{r}
#Using kruskal-wallace test
kruskal.test(as.numeric(meta.norare.abund$Tube.Color) ~ gfe_20percent, data=meta.norare.abund)
#p-value = 0.4822
#No significant difference
pairwise.wilcox.test(as.numeric(meta.norare.abund$Tube.Color), meta.norare.abund$gfe_20percent, p.adjust.method="fdr")
#No difference

kruskal.test(as.numeric(meta.norare.abund$Tube.Color) ~ rfi_Efficiency, data=meta.norare.abund)
#p-value = 0.7042
#No significant difference
pairwise.wilcox.test(as.numeric(meta.norare.abund$Tube.Color), meta.norare.abund$rfi_Efficiency, p.adjust.method="fdr")
#No difference

```


##Testing diff abundance for GFE
```{r, Differential abundance testing with DESeq2}


#Using the GFE Efficiency Label
sample_data(physeq.norare.abund)$gfe_20percent <- as.factor(sample_data(physeq.norare.abund)$gfe_20percent)

ds = phyloseq_to_deseq2(physeq.norare.abund, ~ gfe_20percent)

ds = DESeq(ds)

alpha = 0.05
res = results(ds, contrast=c("gfe_20percent", "HE", "LE"), alpha=alpha)
res = res[order(res$padj, na.last=NA), ]
res_sig = res[(res$padj < alpha), ]
res_sig$Comp <- "GFE"
res_sig

res_sig = cbind(as(res_sig, "data.frame"), as(tax_table(physeq.norare.abund)[rownames(res_sig), ], "matrix"))
ggplot(res_sig, aes(x=Genus, y=log2FoldChange, color=Phylum)) +
    geom_jitter(size=3, width = 0.2) +
    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
    ggtitle("DESeq2 Second Lactation GFE High vs Low")

par(mfrow=c(1,1))

```

##Testing diff abundance for RFI
```{r}
#Using the RFI Efficiency Label
sample_data(physeq.norare.abund)$rfi_Efficiency <- as.factor(sample_data(physeq.norare.abund)$rfi_Efficiency)

ds.rfi = phyloseq_to_deseq2(physeq.norare.abund, ~ rfi_Efficiency)
ds.rfi = DESeq(ds.rfi)

alpha = 0.05
res.rfi = results(ds.rfi, contrast=c("rfi_Efficiency", "HE", "LE"), alpha=alpha)
res.rfi = res.rfi[order(res.rfi$padj, na.last=NA), ]
res_sig.rfi = res.rfi[(res.rfi$padj < alpha), ]
res_sig.rfi
res_sig.rfi$Comp <- "RFI"

res_sig.rfi = cbind(as(res_sig.rfi, "data.frame"), as(tax_table(physeq.norare.abund)[rownames(res_sig.rfi), ], "matrix"))
ggplot2::ggplot(res_sig.rfi, aes(x=Genus, y=log2FoldChange, color=Phylum)) +
    geom_jitter(size=3, width = 0.2) +
    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
    ggtitle("DESeq2 Second Lactation RFI High vs Low")


#We have ASVs of interest. Will take them and compare relative abundance of them.

```


```{r}
DESeq_Results <- rbind(res_sig, res_sig.rfi)

write.csv(DESeq_Results, file="DESeq2 Results for GFE and RFI Groups.csv")
```


#Linear Modelling with CLR
```{r}
meta.second <- as.matrix(sample_data(physeq))
meta.second <- as.data.frame(meta.second)

#Applying PCA to the centered log-ratio (CLR) transformed sequence counts
ps_clr <- microbiome::transform(physeq, "clr")

#PCA via phyloseq
ord_clr <- phyloseq::ordinate(ps_clr, "RDA")

#Plot scree plot
phyloseq::plot_scree(ord_clr) +
  geom_bar(stat="identity", fill="blue") +
  labs(x = "\nAxis", y = "Proportion of Variance\n")

#Examine eigenvalues and % proportional variance explained
head(ord_clr$CA$eig)

sapply(ord_clr$CA$eig[1:5], function(x) x / sum(ord_clr$CA$eig))     

#Scale axes and plot ordination
clr1 <- ord_clr$CA$eig[1] / sum(ord_clr$CA$eig)
clr2 <- ord_clr$CA$eig[2] / sum(ord_clr$CA$eig)
z_second_rfi <- phyloseq::plot_ordination(physeq, ord_clr, type="samples", color="rfi_Efficiency") + 
  geom_point(size = 2) +
  coord_fixed() +  
  ylim(-20,20) +
  xlim(-20,20) +
  stat_ellipse(aes(group = rfi_Efficiency), linetype = 2) +
  ggtitle("Second Lactation PCA by RFI") +
  guides(color = guide_legend(title = "RFI Group")) +
  scale_color_manual(values = okabe_ito_colors)

z_second_rfi

z_second_gfe <- phyloseq::plot_ordination(physeq, ord_clr, type="samples", color="gfe_20percent") + 
  geom_point(size = 2) +
  coord_fixed() +  
  ylim(-20,20) +
  xlim(-20,20) +
  stat_ellipse(aes(group = gfe_20percent), linetype = 2) +
  ggtitle("Second Lactation PCA by GFE") +
  guides(color = guide_legend(title = "GFE Group")) +
  scale_color_manual(values = okabe_ito_colors)

z_second_gfe


ordination_plot <- plot_ordination(physeq, ord_clr, type="samples", color="rfi_Efficiency")
plot_data <- as.data.frame(ordination_plot$data)


#Test for differences between groups
#Generate distance matrix
clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean") 
#ADONIS test
vegan::adonis2(clr_dist_matrix ~ sample_data(ps_clr)$rfi_Efficiency)
#No significance
vegan::adonis2(clr_dist_matrix ~ sample_data(ps_clr)$gfe_20percent)
#Trending p = 0.093


#Pull out this PC1 column, combine with the GFE_Efficiency column, and then plot!
plot_data <- plot_data[order(row.names(plot_data)),]
meta.second.2 <- meta.second[order(row.names(meta.second)),]

plot_data_2 <- as.data.frame(plot_data[,1:2])
row.names(plot_data_2) <- row.names(plot_data)
colnames(plot_data_2)[1] <- "PC1"
colnames(plot_data_2)[2] <- "PC2"

meta.second.3 <- merge(meta.second.2, plot_data_2, by="row.names")
row.names(meta.second.3) <- meta.second.3$Row.names


meta.second.3$rfi_recent <- as.numeric(meta.second.3$rfi_recent)
meta.second.3$GFE <- as.numeric(meta.second.3$GFE)


##Linear modelling of GFE/RFI against PC1 axis for ALL ANIMALS
GFE.pc1.2 <- ggplot(meta.second.3, aes(GFE, PC1), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    xlab("Gross Feed Efficiency") +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("GFE x PC1")

GFE.pc2.2 <- ggplot(meta.second.3, aes(GFE, PC2), geom_smooth(method="lm", se=FALSE)) +
    geom_point() +
    xlab("Gross Feed Efficiency") +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("GFE x PC2")




RFI.pc1.2 <- ggplot(meta.second.3, aes(rfi_recent, PC1)) +
    geom_point() +
    xlab("Residual Feed Intake") +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("RFI x PC1")

RFI.pc2.2 <- ggplot(meta.second.3, aes(rfi_recent, PC2)) +
    geom_point() +
    xlab("Residual Feed Intake") +
    geom_smooth(method="lm", se=FALSE) +
    ggtitle("RFI x PC2")


lm_fit_GFE.PC1 <- lm(GFE ~ PC1, data=meta.second.3)
summary(lm_fit_GFE.PC1)
#Not significant

lm_fit_RFI.PC1 <- lm(rfi_recent ~ PC1, data=meta.second.3)
summary(lm_fit_RFI.PC1)
#Significant!



lm_fit_GFE.PC2 <- lm(GFE ~ PC2, data=meta.second.3)
summary(lm_fit_GFE.PC2)
#Not significant

lm_fit_RFI.PC2 <- lm(rfi_recent ~ PC2, data=meta.second.3)
summary(lm_fit_RFI.PC2)
#Not significant
 
RFI.pc1.2

grid.arrange(GFE.pc1.2, GFE.pc2.2, RFI.pc1.2, RFI.pc2.2, top = "Second Lactation", ncol = 2)

```

























